A Plasma Torus Around a Young Low-Mass Star

### 2025-03-14 paper/ms.tex
update

\begin{abstract}
\normalfont
Approximately one percent of red dwarfs younger than 100 million years
show structured, periodic optical light curves suggestive of
transiting opaque material corotating with the star
\cite{Rebull2016,Stauffer2017,Rebull2018,Bouma2024}.  The
composition, origin, and even the existence of this material are
uncertain. The main alternative hypothesis is that these complex
periodic variables (CPVs) are explained by complex distributions of
starspots or faculae distributed across the stellar surfaces
\cite{Koen2021}.  Here, we present time-series spectroscopy and
photometry of a rapidly-rotating ($P$=3.9\,hr)
CPV, TIC~141146667. The spectra show coherent sinusoidal Balmer
emission at twice to four times the star's equatorial velocity,
demonstrating the presence of extended clumps of circumstellar
plasma --- a plasma torus.  Given that long-lived condensations of
cool ($10^4$ K) plasma can persist in the hot ($10^6$ K) coronae of
stars with a wide range of masses
\cite{CollierCameron1989,Townsend2005,Dunstone2006,Petit2013,Waugh2022,Daley-Yates2024},
these data support the idea that around the lowest-mass stars, such
condensations can be optically thick; the exact origin of the
material and its microphysical opacity remain unclear.
\end{abstract}

@@

The two leading hypotheses to explain the CPVs are either that
transiting clumps of circumstellar material corotate with the star
\cite{Stauffer2017,Gunther2022,Bouma2024}, or that these stars
represent an extreme in naturally-occurring distributions of starspots
or faculae \cite{Koen2021}.  Currently, the main argument against a
starspot-only explanation invokes the timescales and amplitudes of the
sharpest photometric features.  However, no independent evidence has
yet been acquired for the presence of circumstellar material around these
objects.  
Transiting circumstellar clumps would geometrically imply an intrinsic
occurrence a few to ten times the observed rate.  The question of
whether circumstellar clumps exist in these systems could therefore be
relevnat for 10-30\% of M dwarfs during their early lives.

@@

Our Keck/HIRES observations are the first reported time-series
spectra of a CPV, and they show that corotating circumstellar plasma clumps 
are present in at least one such star.
Characteristic densities and masses of these
%TODO check; quote N_H instead...
clumps are $n \sim 10^{10}$\,cm$^{-3}$ and $M \sim 10^{14}$\,kg (see
Supplementary Methods Section~\ref{subsec:model}), a similar density
to solar prominences, but ten to one hundred times more massive.  This
observation rules out a ``starspot-only'' origin scenario for CPVs,
\cite{Koen2021} since such scenarios have no means of explaining
spectroscopic emission beyond the stellar disk.  Similarly, scenarios
in which the circumstellar material is made only of dust are also
ruled out.  While dust may be present, to explain the H$\alpha$
emission the circumstellar clumps must include plasma with a
significant population of hydrogen atoms in the $n$=3 excited state.
While this plasma is undoubtedly sculpted by the star's magnetic
field, it could plausibly originate from three sites: the star, an old
and undetected disk, or outgassing rocky bodies.  This latter
possibility would render CPVs as extrasolar analogs of the Jupiter-Io
plasma torus (e.g.~Ref.~\cite{Bagenal1981}).

@@

The role of the star's magnetic field could be better understood
through new observations, and new theory.  From the theoretical
perspective, there is an urgent need for rigid-field
(magneto)-hydrodynamic modeling to go beyond previous work
\cite{Townsend2005,Townsend2008} and to explore the effects of
non-dipolar field contributions.
In particular, dynamo simulations of fully-convective M dwarfs have
suggested that global-scale mean fields might be
confined to a single hemisphere \cite{Brown2020}; such fields would
yield accumulations of circumstellar material quite different
from those that have previously been explored \cite{Townsend2008}.
Observationally, spectropolarimetry has the potential to
assess both the field strength and topology.  An independent
probe could also be connected the recent work \cite{Kaur2024} showing
that CPVs are variable radio emitters with emission components
that can be both persistent, and also short-lived and highly
polarized.  In particular, detecting radio emission produced by an electron
cyclotron maser instability would provide a measurement of the field
strength at the site of the emitting region.

@@

For TIC~141146667, we measured the radial velocities using regions
near the K~I (7700\,\AA) resonance line and three TiO bandheads
(5160\,\AA, 5450\,\AA, and 5600\,\AA).  We
selected these regions because they provided the best matches between the
synthetic and observed spectra.  We then averaged the resulting redshift measurements
over each order to produce the final measurement.  We used
the scatter of resulting velocity measurements between orders to assign
the RV uncertainty at each epoch.  The uncertainty-weighted mean
systemic velocity over all epochs on UT 2024-02-17 was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are given in Table~\ref{tab:rv}.

@@

Figure~\ref{fig:camd} shows that TIC~141146667 falls between
the USco and Pleiades sequences, while nearly overlapping with
IC~2602.  However, the precision of the implied age
is set by the intrinsic scatter of these calibration
sequences; the most luminous stars in the Pleiades of the same color
have a similar absolute magnitude as TIC~141146667.
Previous work \cite{Stauffer2021} has also noted that in the Gaia
passbands, CPVs tend to be photometrically redder and more luminous
than single stars in any given cluster, similar to other rapid
rotators.
While this effect complicates any attempt at age inference based on
the Gaia photometry, it suggests that the Pleiades may be a better
comparison population than IC~2602.  Broadly, we take the star's
isochrone location to suggest age bounds $t_{\rm
CAMD}$$\sim$40-150\,Myr.

%  and applied the same reddening corrections as in \cite{Bouma2022}.
%  To measure the isochrone age of TIC~141146667, we then followed the 
%  empirical approach proposed by \cite{Gagne2020} and
%  implemented in \cite{Bouma2022}.
%  The idea is to estimate the star (or star cluster's) age by
%  interpolating between the observed isochrones of star clusters with
%  ages that have been otherwise calibrated, for instance using
%  the lithium depletion-boundary method.
%  While this circumvents known modelling issues in molecular opacities
%  and starspot coverage for young M dwarfs,
%  the ages returned by this procedure depend on the ages
%  assumed for each reference cluster.
%  In this case, the most important reference is UCL, since TIC~141146667
%  overlaps directly on its isochrone on the CAMD.
%  We adopted the 16~Myr age for UCL from \cite{PecautMamajek2016}, which
%  is consistent with ages determined by more recent sub-clustering
%  analyses \cite{Ratzenbock2023} for the largest sub-populations.
%  This in turns yields an isochronal age for TIC~141146667 of
%  $t_{\rm iso}$=$16^{+19}_{-6}$\,Myr.
%  Qualitatively, this range of uncertainty is consistent with no stars
%  in the $\approx$40\,Myr IC~2602 cluster with similar redness as
%  TIC~141146667 being equally bright.
%FIXME: update discussion following consideration of anomalous CPV
% brightnesses and colors...

@@

Desite uncertainties in the details, what can be stated with
confidence is that lithium is not abundant in the spectrum of
TIC~141146667.  Figure~\ref{fig:liew_population} compares our upper
limit against the lithium equivalent width measurements reported by
\cite{Jeffries2023} based on the Gaia-ESO spectroscopic survey.  If
the star were $\lesssim$20\,Myr old, given its temperature we would
expect to see lithium in abundance ($>$400\,m\AA).  Since we do not,
we conclude that $t_{\rm Li,emp}$$\gtrsim$20\,Myr.  The
\cite{Feiden2016} isochrones provide a point for theoretical
comparison, suggesting that since $M_{\rm K}$$=$6.67\,mag, $t_{\rm
Li,th}$$\gtrsim$35\,Myr is the theoretical age at which complete
depletion occurs in a star with this luminosity (see e.g.~Figure 7
from \cite{Wood2023}).

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf3.pdf}
  \caption{Dereddened Gaia DR3 color vs.~absolute magnitude diagram of
  TIC~141146667 and comparison samples. 
  TIC~141146667 is on the pre-main sequence; stars with the same color
  on the main sequence are $\approx$1.5\,magnitudes
  fainter.  The star's location in this diagram broadly suggests an
  age of 40-150\,Myr.  }
  \label{fig:camd}
\end{figure}

@@

{\it Age: Summary}---The main indicators for the youth of
TIC~141146667 are {\it i)} that it is a complex periodic variable, and
{\it ii)} that it is 1.5 magnitudes brighter than main sequence stars
of the same color, while showing no indicators for binarity.  The
first item constitutes evidence for youth since all known CPVs are
$\lesssim$200\,Myr \cite{Bouma2024}.  The reference sequences for the
Pleiades (112\,Myr, \cite{Dahm2015}) show a few stars of equal
luminosity and the same temperature, suggesting a photometric
isochronal age upper limit $\lesssim$150\,Myr.  The weak lithium
absorption suggests an age of at least 20\,Myr based on an empirical
comparison using Gaia-ESO spectra, or at least 35\,Myr based on the
\cite{Feiden2016} isochrones.  These considerations yield our adopted
age range of 40-150\,Myr.

%%%%%%%%
%\texttt{astroARIADNE} compares the measured broadband flux
%measurements against pre-computed model grids, and by default fits for
%six parameters: $\{ T_{\rm eff}, R_\star, A_{\rm V}, \log g, [{\rm
%Fe/H}], d \}$.  The distance  prior is drawn from
%\citet{2021AJ....161..147B}.  The surface gravity and metallicity are
%generally unconstrained.  Given our selection criteria for the stars,
%we assumed the following priors for the temperature, stellar size, and
%extinction:
%\begin{align}
%  T_{\rm eff} / {\rm K}    &\sim \mathcal{U}(2000, 8000), \\
%  R_\star / R_\odot  &\sim \mathcal{U}(0.1, 1.5), \\
%  A_{\rm V} / {\rm mag}    &\sim \mathcal{U}(0, 0.2),
%\end{align}
%for \deleted{$\mathcal{N}$ the Gaussian and }$\mathcal{U}$ the uniform
%distribution\deleted{s, and $\mathcal{T}_{\rm N}(\mu, \sigma, a, b)$ a
%truncated normal distribution with mean $\mu$, standard deviation
%$\sigma$, and lower and upper bounds $a$ and $b$}.  We validated our
%chosen upper bound on $A_{\rm V}$ using a 2MASS color-color diagram.
%Finally, using \texttt{Dynesty} \citep{2020MNRAS.493.3132S}, we
%sampled the posterior probability assuming the default Gaussian
%likelihood, and set a stopping threshold of ${\rm d}\log \mathcal{Z} <
%0.01$, where $\mathcal{Z}$ denotes the evidence.

{\it SED: No evidence for disk}
...

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (\kms) & & \\
    $V$ (\kms) & & \\
    $W$ (\kms) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & This work\\
    $i_\star$($^\circ$) & $>$63 & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
    %$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & 0.188 $\pm$ 0.012  & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}

>>>>

\begin{abstract}
\normalfont
Approximately one percent of red dwarfs younger than 100 million years
show structured, periodic optical light curves suggestive of
transiting opaque material that corotates with the star
\cite{Rebull2016,Stauffer2017,Rebull2018,Bouma2024}.  The
composition, origin, and even the existence of this material are
uncertain. The main alternative hypothesis is that these complex
periodic variables (CPVs) are explained by complex distributions of
bright or dark regions on the stellar surfaces
\cite{Koen2021}.  Here, we present time-series spectroscopy and
photometry of a rapidly-rotating ($P$=3.9\,hr)
CPV, TIC~141146667. The spectra show coherent sinusoidal Balmer
emission at twice to four times the star's equatorial velocity,
demonstrating the presence of extended clumps of circumstellar
plasma --- a plasma torus.  Given that long-lived condensations of
cool ($10^4$ K) plasma can persist in the hot ($10^6$ K) coronae of
stars with a wide range of masses
\cite{CollierCameron1989,Townsend2005,Dunstone2006,Petit2013,Waugh2022,Daley-Yates2024},
these data support the idea that around the lowest-mass stars, such
condensations can be optically thick; the exact origin of the
material and its microphysical opacity remain unclear.
\end{abstract}

@@

The two leading hypotheses to explain the CPVs are either that
transiting clumps of circumstellar material corotate with the star
\cite{Stauffer2017,Gunther2022,Bouma2024}, or that these stars
represent an extreme in naturally-occurring distributions of starspots
or faculae \cite{Koen2021}.  Currently, the main argument against a
starspot-only explanation invokes the timescales and amplitudes of the
sharpest photometric features.  However, no independent evidence has
yet been acquired for the presence of any circumstellar material.  
Geometrically, transiting clumps would imply an occurrence
a few to at most ten times the observed rate.
Such clumps could therefore exist around 10-30\% of M dwarfs during
their early lives.

@@

Our Keck/HIRES observations are the first reported time-series
spectra of a CPV, and they show that corotating circumstellar plasma clumps 
are present in at least one such star.
Characteristic densities and masses of these
%TODO check; quote N_H instead...
clumps are $n \sim 10^{10}$\,cm$^{-3}$ and $M \sim 10^{14}$\,kg (see
Supplementary Methods Section~\ref{subsec:model}), a similar density
to solar prominences, but ten to one hundred times more massive.  This
observation rules out a ``starspot-only'' origin scenario for CPVs,
\cite{Koen2021} since such scenarios have no means of explaining
spectroscopic emission beyond the stellar disk.  Similarly, scenarios
in which the circumstellar material is made only of dust are also
ruled out.  While dust may be present, to explain the H$\alpha$
emission the clumps must include plasma with a
significant population of hydrogen atoms in the $n$=3 excited state.
While this plasma is undoubtedly sculpted by the star's magnetic
field, it could plausibly originate from three sites: the star, an old
and undetected disk, or outgassing rocky bodies.  This latter
possibility would render CPVs as extrasolar analogs of the Jupiter-Io
plasma torus (e.g.~Ref.~\cite{Bagenal1981}).

@@

The role of the star's magnetic field could be better understood
through new observations, and new theory.  From the theoretical
perspective, there is an urgent need for rigid-field
(magneto)-hydrodynamic modeling to go beyond previous work
\cite{Townsend2005,Townsend2008} and to explore the effects of
non-dipolar field contributions.
In particular, dynamo simulations of fully-convective M dwarfs have
suggested that global-scale mean fields might be
confined to a single hemisphere \cite{Brown2020}; such fields could
yield accumulation surfaces quite different
from those that have previously been explored \cite{Townsend2008}.
Observationally, spectropolarimetry has the potential to
assess both the field strength and topology.  An independent
probe could also be connected the recent work \cite{Kaur2024} showing
that CPVs are variable radio emitters with emission components
that can be both persistent, and also short-lived and highly
polarized.  In particular, detecting radio emission produced by an electron
cyclotron maser instability would provide a measurement of the field
strength at the site of the emitting region.

@@

For TIC~141146667, we measured the radial velocities using regions
near the K~I (7700\,\AA) resonance line and three TiO bandheads
(5160\,\AA, 5450\,\AA, and 5600\,\AA).  We
selected these regions because they provided the best matches between the
synthetic and observed spectra.  We then averaged the resulting redshift measurements
over each order to produce the final measurement.  We used
the scatter of resulting velocity measurements between orders to assign
the RV uncertainty at each epoch.  The uncertainty-weighted mean
systemic velocity over all epochs on 17~February~2024 (UT) was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are given in Table~\ref{tab:rv}.

@@

Figure~\ref{fig:camd} shows that TIC~141146667 falls between the USco
and Pleiades sequences, while nearly overlapping with IC~2602.  However,
the precision of the implied age is set by the intrinsic scatter of
these calibration sequences; the most luminous stars in the Pleiades of
the same color have a similar absolute magnitude as TIC~141146667.
Previous work \cite{Stauffer2021} has also noted that in the Gaia
passbands, CPVs tend to be photometrically redder and more luminous than
single stars in any given cluster, similar to other rapid rotators.
While this effect complicates any attempt at age inference based on the
Gaia photometry, it suggests that the Pleiades may be a better
comparison population than IC~2602.  Broadly, we take the star's
location in the color--absolute magnitude diagram to suggest age bounds
$t_{\rm CAMD}$$\sim$30-150\,Myr.

@@

Despite uncertainties in the details, what can be stated with confidence
is that lithium is not abundant in the spectrum of TIC~141146667.
Figure~\ref{fig:liew_population} compares our upper limit against the
equivalent width measurements reported by \cite{Jeffries2023}
based on the Gaia-ESO spectroscopic survey.  If the star were
$\lesssim$20\,Myr old, at its temperature we would expect to see
lithium in abundance ($>$400\,m\AA).  Since we do not, we can set an
empirical bound on the lithium-derived age of $t_{\rm
Li,emp}$$\gtrsim$20\,Myr.  The \cite{Feiden2016} lithium isochrones
provide a point for theoretical comparison, and suggest that since
$M_{\rm K}$$=$6.67\,mag, $t_{\rm Li,th}$$\gtrsim$35\,Myr is the
theoretical age at which complete depletion occurs in a star with this
luminosity (see e.g.~Figure 7 from \cite{Wood2023}).

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf3.pdf}
  \caption{Dereddened Gaia DR3 color vs.~absolute magnitude diagram of
  TIC~141146667 and comparison samples. 
  TIC~141146667 is on the pre-main sequence; stars with the same color
  on the main sequence are $\approx$1.5\,magnitudes
  fainter.  The star's location in this diagram broadly suggests an
  age of 30-150\,Myr.  }
  \label{fig:camd}
\end{figure}

@@

{\it Age: Summary}---The main indicators for the youth of
TIC~141146667 are {\it i)} that it is a complex periodic variable, and
{\it ii)} that it is 1.5 magnitudes brighter than main sequence stars
of the same color, while showing no indicators for binarity.  The
first item constitutes evidence for youth since all known CPVs are
$\lesssim$200\,Myr \cite{Bouma2024}.  The reference sequences for the
Pleiades (112\,Myr, \cite{Dahm2015}) show a few stars of equal
luminosity and the same temperature, suggesting a photometric
isochronal age upper limit $\lesssim$150\,Myr.  The weak lithium
absorption suggests an age of at least 20\,Myr based on an empirical
comparison using Gaia-ESO spectra, or at least 35\,Myr based on the
\cite{Feiden2016} isochrones.  These considerations yield our adopted
age range of 150\,Myr.

{\it SED: No evidence for disk}---Figure~\ref{fig:sed} shows the 
...

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry \& Radial Velocity} \\ 
    \hline
    $\alpha_{2000}$ & 11:05:15.09   & Gaia\ DR3 \\
    $\delta_{2000}$ & +59 15 05.57  & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    RV (\kms)                     & 0.61 $\pm$ 1.47     & HIRES \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag)                    & 13.283 $\pm$ 0.010 & TIC8\     \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    %\hline
    %\multicolumn{3}{c}{Kinematics and Position} \\
    %\hline
    %$RV$ (\kms) & 0.61 $\pm$ 1.47 & HIRES \\
    %$U$ (\kms)  & & \\
    %$V$ (\kms)  & & \\
    %$W$ (\kms)  & & \\
    %$X$ (pc)    & -28.4 & \\
    %$Y$ (pc)    &  19.8 & \\
    %$Z$ (pc)    &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024} SED fit\\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} SED fit \\
    $P_{\rm rot}$ (hours) & $3.930\pm 0.001$ & TESS \\ 
    $v_{\rm eq}$ (\kms)  &  130$\pm$4  & Derived \\
    $v_{\rm eq} \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & HIRES \\
    $i_\star$($^\circ$) & $>$63 & Derived \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    %$F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    %$L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & 0.22 $\pm$ 0.02  & PARSEC \cite{Chen2014}\\
    ${\rm EW}_{\rm Li}$ (m\AA) & $<$107 & HIRES (2$\sigma$)\\
    $t_{\rm CAMD}$ (Myr) & 30-150 &  Gaia DR3 \\
    $t_{\rm Li,emp}$ (Myr) & $>$20 &  HIRES, \cite{Jeffries2023} \\
    $t_{\rm Li,th}$ (Myr) & $>$35 &  HIRES, \cite{Feiden2016} \\
    $t_{\rm adopted}$ (Myr) & 35-150 &  -- \\
    \hline
    \end{tabular}
		\caption{Properties of \starname.  References:
    Gaia DR3 \cite{GaiaDR3}, TESS \cite{Ricker2015},
    TIC8 \cite{Stassun2019}, 2MASS \cite{Skrutskie2006}, ALLWISE
    \cite{Cutri2014}.}
    \label{tab:stparams}
\end{table}
